---
title: "TINS ShinyApp Data Exploration"
author: "Cameryn Brock"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

### Ideas for graphs

- Number of birds banded and recaps per year / total 
- Number of species / year
- Number of birds banded / species banded
- Number of birds banded with colors for species / families? - Need to put in families and orders but should be easy enough
- Capture rate? Number / net hour
- Check any change in weight / wing size over time within species
- Toggle between birds per net hour and total (or show both graphs)

### Ideas for dashboards

- Could do big graph with number of birds banded per year, then you click and see # orders, families, species
- Could drop down menu of species (family, order) and see number banded per year / per some standardized effort and could allow to overlay on top of each other
- Visual spatial data with recaps
- Consider making a web map of Tahoe banded species? 

```{r}

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
library(paletteer)
library(tsibble)

```

```{r}
# Read in and clean up data to combine separate year files to one df

tins_2010 <- read_csv(here::here("RawData", "2010.csv"), 
                      col_types = cols(.default = "c")) %>%  
  select(-c(29, 30, 32)) %>% 
  head(212) 

tins_2011 <- read_csv(here::here("RawData", "2011.csv"), 
                      col_types = cols(.default = "c")) %>% 
  select(-c(29, 30, 32)) 

tins_2012 <- read_csv(here::here("RawData", "2012.csv"), 
                      col_types = cols(.default = "c"))

tins_2013 <- read_csv(here::here("RawData", "2013.csv"), 
                      col_types = cols(.default = "c"))
 
tins_2014 <- read_csv(here::here("RawData", "2014.csv"), 
                      col_types = cols(.default = "c"))
 
tins_2015 <- read_csv(here::here("RawData", "datafixed2015.csv"), 
                      col_types = cols(.default = "c")) %>% 
  select(-c(25:27)) %>% 
  rename("NOTE?" = "Comment")
 
tins_2016 <- read_csv(here::here("RawData", "2016.csv"), 
                      col_types = cols(.default = "c")) %>% 
  select(-c(25:27)) %>% 
  head(778) %>% 
  rename("NOTE?" = "Comment")

tins_2017 <- read_csv(here::here("RawData", "2017.csv"), 
                      col_types = cols(.default = "c")) %>% 
  head(167) %>% 
  rename("NOTE?" = "Comment")
 
tins_2018 <- read_csv(here::here("RawData", "2018.csv"), 
                      col_types = cols(.default = "c")) %>% 
  head(200) %>% 
  rename("NOTE?" = "Comment")
 
tins_2019 <- read_csv(here::here("RawData", "2019.csv"), 
                      col_types = cols(.default = "c")) %>% 
  head(29) %>% 
  rename("NOTE?" = "Comment")

# Read in and clean up net hours data

nethours_daily <- read_csv(here::here("RawData", "Nethours2010.csv")) %>% 
  clean_names() %>% 
  mutate(date = mdy(date))

```

```{r} 
# bind all years

tins <- bind_rows(tins_2010,
                  tins_2011,
                  tins_2012,
                  tins_2013,
                  tins_2014,
                  tins_2015,
                  tins_2016,
                  tins_2017,
                  tins_2018,
                  tins_2019)

# Tidy tins

tins_tidy <- tins %>% 
  clean_names() %>% 
  mutate(wing = as.numeric(wing),
         wght = as.numeric(wght)) %>% 
  mutate(date = lubridate::mdy(date),
         year = lubridate::year(date),
         month = lubridate::month(date),
         doy = lubridate::yday(date)) %>% 
  filter(!year == "2003") %>% 
  mutate(spec = case_when(spec %in% c("AUWA", "MYWA", "YRWA") ~ "YRWA",
                          spec %in% c("GWCS", "MWCS", "WCSP") ~ "WCSP",
                          T ~ spec)) %>% 
  filter(!spec == "NA")

# Did the case_when to remove subspecies & only include species
# Checked all species codes to ensure no more duplicates or typos
# Some that weren't typical: 
# We have RSFL instead of NOFL (red shafted flicker instead of northern)
# We have WEFL (western flycatcher) - it has since been split into PSFL and COFL
# We have YWAR for yellow warbler, looks like it is sometimes YEWA? 

# Update sites and time of interest
# Add taxonomy and viewer-friendly labels

tins_roi <- tins_tidy %>% 
  filter(loc %in% c("COEA", "BLCA", "POMA")) %>% 
  filter(month %in% c(8, 9, 10)) %>% 
  mutate(family = case_when(spec %in% c("STJA") ~ "Corvidae",
                            spec %in% c("RSFL", "WISA", "DOWO",
                                        "RBSA") ~ "Picidae",
                            spec %in% c("SSHA", "COHA") ~ "Accipitridae",
                            spec %in% c("AMRO", "HETH") ~ "Turdidae",
                            spec %in% c("BHGR", "WETA", "LAZB") ~ "Cardinalidae",
                            spec %in% c("GTTO", "WCSP", "SOSP", 
                                        "SAVS", "LISP", "BRSP",
                                        "ORJU", "SPTO", "GCSP",
                                        "CHSP", "FOSP", "LASP") ~ "Passerellidae",
                            spec %in% c("MOCH") ~ "Paridae",
                            spec %in% c("OSFL", "WIFL", "GRFL", 
                                        "WEWP", "DUFL", "HAFL",
                                        "BLPH", "WEFL") ~ "Tyrannidae",
                            spec %in% c("WAVI", "CAVI") ~ "Vireonidae",
                            spec %in% c("BCHU", "RUHU", "ANHU") ~ "Trochilidae",
                            spec %in% c("OCWA", "YWAR", "YRWA",
                                        "MGWA", "WIWA", "BTYW",
                                        "COYE", "NAWA", "HEWA") ~ "Parulidae", 
                            spec %in% c("PUFI", "PISI", "EVGR",
                                        "LEGO") ~ "Fringillidae",
                            spec %in% c("RCKI", "GCKI") ~ "Regulidae",
                            spec %in% c("RBNU", "PYNU") ~ "Sittidae",
                            spec %in% c("BRCR") ~ "Certhiidae",
                            spec %in% c("BUSH") ~"Aegithalidae",
                            spec %in% c("BHCO", "BUOR") ~ "Icteridae",
                            spec %in% c("MAWR", "HOWR", "BEWR") ~ "Troglodytidae",
                            spec %in% c("BARS") ~ "Hirundinidae",
                            spec %in% c("WISN") ~ "Scolopacidae",
                            spec %in% c("MODO") ~ "Columbidae",
                            spec %in% c("BGGN") ~ "Polioptilidae")) %>% 
  mutate(family_label = case_when(family == "Corvidae" ~ "Corvidae: Crows, Jays, & Magpies",
                                  family == "Picidae" ~ "Picidae: Woodpeckers",
                                  family == "Accipitridae" ~ "Accipitridae: Hawks, Eagles, & Kites",
                                  family == "Turdidae" ~ "Turdidae: Thrushes & Allies",
                                  family == "Cardinalidae" ~ "Cardinalidae: Cardinals & Allies",
                                  family == "Passerellidae" ~ "Passerellidae: Sparrows",
                                  family == "Paridae" ~ "Paridae: Chickadees & Titmice",
                                  family == "Tyrannidae" ~ "Tyrant Flycatchers",
                                  family == "Vireonidae" ~ "Vireonidae: Vireos",
                                  family == "Trochilidae" ~ "Trochilidae: Hummingbirds",
                                  family == "Parulidae" ~ "Parulidae: Warblers",
                                  family == "Fringillidae" ~ "Fringillidae: Finches & Allies",
                                  family == "Regulidae" ~ "Regulidae: Kinglets",
                                  family == "Sittidae" ~ "Sittidae: Nuthatches",
                                  family == "Certhiidae" ~ "Certhiidae: Treecreepers",
                                  family == "Aegithalidae" ~ "Aegithdalidae: Long-tailed Tits",
                                  family == "Icteridae" ~ "Blackbirds & Orioles",
                                  family == "Troglodytidae" ~ "Troglodytidae: Wrens",
                                  family == "Hirundinidae" ~ "Hirundinidae: Swallows",
                                  family == "Scolopacidae" ~ "Sandpipers & Allies",
                                  family == "Columbidae" ~ "Columbidae: Pigeons & Doves",
                                  family == "Polioptilidae" ~ "Polioptilidae: Gnatcatchers")) %>% 
  mutate(spec_label = case_when(spec == "STJA" ~ "Steller's Jay",
                                spec == "RSFL" ~ "Northern Flicker",
                                spec == "WISA" ~ "Williamson's Sapsucker",
                                spec == "DOWO" ~ "Downy Woodpecker",
                                spec == "RBSA" ~ "Red-breasted Sapsucker",
                                spec == "SSHA" ~ "Sharp-shinned Hawk",
                                spec == "COHA" ~ "Cooper's Hawk",
                                spec == "AMRO" ~ "American Robin",
                                spec == "HETH" ~ "Hermit Thrush",
                                spec == "BHGR" ~ "Black-headed Grosbeak",
                                spec == "WETA" ~ "Western Tanager",
                                spec == "LAZB" ~ "Lazuli Bunting",
                                spec == "GTTO" ~ "Green-tailed Towhee",
                                spec == "WCSP" ~ "White-crowned Sparrow",
                                spec == "SOSP" ~ "Song Sparrow",
                                spec == "SAVS" ~ "Savannah Sparrow",
                                spec == "LISP" ~ "Lincoln's Sparrow",
                                spec == "BRSP" ~ "Brewer's Sparrow",
                                spec == "ORJU" ~ "Dark-eyed Junco",
                                spec == "SPTO" ~ "Spotted Towhee",
                                spec == "GCSP" ~ "Golden-crowned Sparrow",
                                spec == "CHSP" ~ "Chipping Sparrow",
                                spec == "FOSP" ~ "Fox Sparrow",
                                spec == "LASP" ~ "Lark Sparrow",
                                spec == "MOCH" ~ "Mountain Chickadee",
                                spec == "OSFL" ~ "Olive-sided Flycatcher",
                                spec == "WIFL" ~ "Willow Flycatcher",
                                spec == "GRFL" ~ "Gray Flycatcher",
                                spec == "WEWP" ~ "Western Wood-Pewee",
                                spec == "DUFL" ~ "Dusky Flycatcher",
                                spec == "HAFL" ~ "Hammond's Flycatcher",
                                spec == "BLPH" ~ "Black Phoebe",
                                spec == "WEFL" ~ "Pacific-slope Flycatcher", # Ask Will if this is ok
                                spec == "WAVI" ~ "Warbling Vireo",
                                spec == "CAVI" ~ "Cassin's Vireo",
                                spec == "BCHU" ~ "Black-chinned Hummingbird",
                                spec == "RUHU" ~ "Rufous Hummingbird",
                                spec == "ANHU" ~ "Anna's Hummingbird",
                                spec == "OCWA" ~ "Orange-crowned Warbler",
                                spec == "YWAR" ~ "Yellow Warbler",
                                spec == "YRWA" ~ "Yellow-rumped Warbler",
                                spec == "MGWA" ~ "MacGillivray's Warbler",
                                spec == "WIWA" ~ "Wilson's Warbler",
                                spec == "BTYW" ~ "Black-throated Gray Warbler",
                                spec == "COYE" ~ "Common Yellowthroat",
                                spec == "NAWA" ~ "Nashville Warbler",
                                spec == "HEWA" ~ "Hermit Warbler",
                                spec == "PUFI" ~ "Purple Finch",
                                spec == "PISI" ~ "Pine Siskin",
                                spec == "EVGR" ~ "Evening Grosbeak",
                                spec == "LEGO" ~ "Lesser Goldfinch",
                                spec == "RCKI" ~ "Ruby-crowned Kinglet",
                                spec == "GCKI" ~ "Golden-crowned Kinglet",
                                spec == "RBNU" ~ "Red-breasted Nuthatch",
                                spec == "PYNU" ~ "Pygmy Nuthatch",
                                spec == "BRCR" ~ "Brown Creeper",
                                spec == "BUSH" ~ "Bushtit",
                                spec == "BHCO" ~ "Brown-headed Cowbird",
                                spec == "BUOR" ~ "Bullock's Oriole",
                                spec == "MAWR" ~ "Marsh Wren",
                                spec == "HOWR" ~ "House Wren",
                                spec == "BEWR" ~ "Bewick's Wren",
                                spec == "BARS" ~ "Barn Swallow",
                                spec == "WISN" ~ "Wilson's Snipe",
                                spec == "MODO" ~ "Mourning Dove",
                                spec == "BGGN" ~ "Blue-gray Gnatcatcher")) %>% 
  mutate(year_week = yearweek(date)) %>% 
  select(family, family_label, spec, spec_label, date, loc, year_week, year, month, doy, note_2)

nethours_weekly <- nethours_daily %>% 
  mutate(month = month(date)) %>% 
  filter(month %in% c(8, 9, 10)) %>% 
  mutate(year_week = yearweek(date)) %>% 
  group_by(year_week) %>% 
  summarize(nethours = sum(total_nethours, na.rm = TRUE))

tins_nethours <- tins_roi %>% 
  group_by(family, family_label, spec, spec_label, year, year_week) %>% 
  summarize(count = n()) %>% 
  inner_join(nethours_weekly, by = "year_week") %>% 
  mutate(count_per_hour = count / nethours)

```


```{r, fig.height = 7, fig.width = 5} 

ggplot(tins_nethours, aes(x = year_week,
                          y = count_per_hour)) + 
  geom_point(aes(color = family),
             alpha = 0.75,
             show.legend = FALSE) +
  geom_smooth(size = 0.5,
              color = "gray60",
              se = FALSE) +
  facet_wrap(~family, 
             scales = "free_y",
             nrow = 8) + 
  theme_minimal() + 
  labs(x = "Year week",
       y = "Count per net hour")

```


```{r, fig.height = 10, fig.width = 5} 

ggplot(tins_nethours, aes(x = year_week,
                          y = count_per_hour)) + 
  geom_point(aes(color = family),
             alpha = 0.75,
             show.legend = FALSE) +
  geom_smooth(size = 0.5,
              color = "gray60",
              se = FALSE) +
  facet_wrap(~spec_label,
             scales = "free_y",
             nrow = 14) + 
  theme_minimal() + 
  labs(x = "Year week",
       y = "Count per net hour")

```


```{r}

# Summarize data

spec_summary <- tins_roi %>% 
  group_by(year, spec) %>% 
  summarize(count = n())

stats_table <- tins_roi %>% 
  group_by(spec) %>% 
  summarize(birds_banded = n()) %>% 
  drop_na()

stats_sum <- stats_table %>% 
  summarize(birds_banded = sum(birds_banded))

```

```{r}

# Try species per net hour per day day / year

spec_day <- tins_roi %>% 
  group_by(date, spec) %>% 
  summarize(no_spec = n()) %>% 
  group_by(date) %>% 
  summarize(no_spec = n())

spec_month <- tins_roi %>% 
  group_by(year, month, spec) %>% 
  summarize(no_spec = n()) %>% 
  group_by(year, month) %>% 
  summarize(no_spec = n())

ind_day <- tins_roi %>% 
  group_by(date) %>% 
  summarize(no_ind = n())

ind_month <- tins_roi %>% 
  group_by(year, month) %>% 
  summarize(no_ind = n())

daily_per_hour <- 
  bind_rows(ind_day, spec_day, nethours_daily) %>% 
  select(date, no_ind, no_spec, total_nethours) %>% 
  group_by(date) %>% 
  summarize(no_ind = sum(no_ind, na.rm = TRUE),
            no_spec = sum(no_spec, na.rm = TRUE),
            nethours = sum(total_nethours, na.rm = TRUE)) %>% 
  filter(!nethours == 0) %>% 
  mutate(spec_per_hour = no_spec / nethours,
         ind_per_hour = no_ind / nethours) %>%
  mutate(year = year(date),
         month = month(date, label = TRUE),
         doy = lubridate::yday(date))

monthly_per_hour <- 
  bind_rows(ind_month, spec_month, nethours_monthly) %>% 
  select(month, no_ind, no_spec, nethours) %>% 
  group_by(year, month) %>% 
  summarize(no_ind = sum(no_ind, na.rm = TRUE),
            no_spec = sum(no_spec, na.rm = TRUE),
            nethours = sum(nethours, na.rm = TRUE)) %>% 
  filter(!nethours == 0) %>% 
  mutate(spec_per_hour = no_spec / nethours,
         ind_per_hour = no_ind / nethours)

ggplot(monthly_per_hour, aes(x = month,
                          y = spec_per_hour,
                          group = year)) +
  geom_line(aes(color = year)) + 
  scale_color_paletteer_c("grDevices::TealRose") + 
  theme_minimal() +
  labs(x = "Month",
       y = "Average species per net hour",
       color = "Year")

ggplot(monthly_per_hour, aes(x = month,
                          y = ind_per_hour,
                          group = year)) +
  geom_line(aes(color = year)) + 
  scale_color_paletteer_c("grDevices::TealRose") + 
  theme_minimal()+
  labs(x = "Month",
       y = "Average individuals per net hour",
       color = "Year")

ggplot(daily_per_hour, aes(x = doy,
                          y = ind_per_hour)) +
  geom_point(aes(color = year),
             size = 3,
             alpha = 0.75) +
  scale_color_paletteer_c("grDevices::TealRose") + 
  theme_minimal() +
  labs(x = "Day of Year",
       y = "Average individuals per net hour",
       color = "Year")

```

```{r}

# Let's look at species specific stuff

# Passerellidae: 

passerellidae <- tins_roi %>% 
  filter(family == "Passerellidae") %>% 
  group_by(spec_label, doy) %>% 
  summarize(count = n())
  

ggplot(passerellidae, aes(x = doy,
                          y = count,
                          color = spec_label)) + 
  geom_point() + 
  theme_minimal()

# Should be count per net hour

```

```{r}

# Let's look at species specific stuff

# Picidae: 

picidae <- tins_roi %>% 
  filter(family == "Picidae") %>% 
  group_by(spec_label, date) %>% 
  summarize(count = n())
  

ggplot(picidae, aes(x = date,
                          y = count,
                          color = spec_label)) + 
  geom_point(alpha = 0.75) + 
  theme_minimal()

# Should be count per net hour

```